---
title: "Women's Political and Economic Participation in Democracies and Autocracies"
author: "Han Na Jun"
description: "GOV50 Final Project"
output:
  distill::distill_article:
    self_contained: false
---


## Introduction

Do women's political participation in parliament positively affect women's economic participation? How do the correlation between these two factors vary across regime types, specifically democratic and authoritarian states? This study draws upon existing literature that investigates authoritarian regimes' incentives for implementing gender quotas, sometimes at higher levels than democratic states. While scholars have argued that authoritarian states do so to gain legitimacy at the domestic and international level, there is relatively little research as to whether such policies addressing women's political representation affect women's empowerment, specifically economic status. 

Accordingly, I seek to investigate the efficacy of women's heightened political participation under such regimes and discover if the policies go beyond mere signaling. I hypothesize that women's political participation in parliament will not have a significant effect on the economic status of women in authoritarian regimes. I also hypothesize that this tendency will be higher for more repressive regimes. This research contributes to the understanding of the substantial effects of women's inclusion in political institutions such as parliament in autocratic states, which is especially pertinent to the appeal for the implementation of gender quotas at the international level.


## Data Selection

My data draws upon the World Development Indicators Database produced by the World Bank and polity score data produced by the Center for Systemic Peace. The unit of analysis is by country and year, and the scope of my analysis is across 217 countries from years 1990 to 2016, considering the rapid increase in gender quotas since the 1990s and the availability of data. My explanatory variable of interest is the 'Proportion of seats held by women in national parliaments (%),' which is a country-level measurement of women's political participation. My outcome variable of interest is the 'Labor force participation rate, female (% of female population ages 15-64),' which is a statistical model estimate by the International Labor Organization. Furthermore, I will compare these results by their relative polity score, which is a country-level coding of authority characteristics of states.

My research is a cross-sectional design that compares the outcome of the treatment, or political participation of women. If I observe little change in the labor force participation of women despite the increase in the number of women in parliament under authoritarian regimes, this would provide support for my hypothesis. If I observe an increase in labor force participation of women along with the increase of women in parliament, the evidence would counter my hypothesis.


```{r}
## Preparing the data

library(tidyverse)
wb_gender <- read_csv("data/wb_gender.csv")
polity <- read_csv("data/polity.csv")

wb_gender <- wb_gender |>
  select(`Country Name`, `Country Code`, `Series Name`, `1990 [YR1990]`:`2016 [YR2016]`)

colnames(wb_gender) <-c("country","ccode","series",1990:2016)

wdi <- wb_gender |>
  gather(key = year, value=value, `1990`:`2016`, factor_key = F) |>
  filter(series == "Proportion of seats held by women in national parliaments (%)" | series == "Labor force participation rate, female (% of female population ages 15-64) (modeled ILO estimate)" | series =="GDP per capita (constant 2015 US$)") |>
  pivot_wider(
    names_from = "series",
    values_from = "value"
  )

colnames(wdi) <- c("country", "ccode", "year", "prop_seats_women", "GDPpc", "labor_female")


wdi$year <- as.numeric(wdi$year)

head(wdi)
```

```{r}
##Joining WDI and Polity data

polity <- polity |>
  pivot_longer(cols = `AFG`:`ZIM`,
    names_to = "ccode",
    values_to = "polity_score"
  ) |>
  mutate(year = as.numeric(year))
polity

join_wdi_pol <- wdi |>
  left_join(polity, by = c("ccode", "year"))
join_wdi_pol

```


Figure 1 is a histogram that summarizes the labor force participation of females across all 217 countries from years 1990 to 2016. The histogram is skewed to the left, indicating that a significant portion of the data lies within the 50% to 75% labor force participation rate of females.


```{r}
## Plotting explanatory variable

plot1 <- wdi |>
  ggplot(aes(x = labor_female)) +
  geom_histogram() +
  labs(title = "Figure 1. Labor Force Participation of Females, 1990 - 2016",
       x = "Female labor force participation rate")
plot1
```



## Results

CORRELATION

To evaluate the relationship between women's political and economic participation, I use both correlation and regression models. I first plot and calculate the correlation between 'Proportion of seats held by women in national parliaments' as my independent variable and 'Female labor force participation rate' as my dependent variable. Figure 2 displays a scatterplot of the two variables in which the direction of the relationship is positive, suggesting that as the proportion of seats held by women in national parliaments increases, the female labor force participation rate also increases. However, the points on the plot are dispersed rather than concentrated, suggesting a weak relationship between the two variables. Most of the points are clustered in the 0-20% and 20-40% range of the proportion of seats held by women in national parliaments, whereas the rate of female participation in the labor force seems to be scattered mostly in the 25-75% range. One outlier sits at over 60% of the proportion of women in seats of parliament. A line of best fit is added to summarize this general trend in the data, which also indicates a meager positive relationship between the two variables.


```{r}
## Plotting dependent and independent variables

plot2 <- join_wdi_pol |>
  ggplot(aes(x = prop_seats_women,
             y = labor_female)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Figure 2. Political and Economic Participation of Females, 1990 - 2016",
       x = "Proportion of seats held by women in national parliaments (%)",
       y = "Female labor force participation rate")
plot2
```


Additionally, I calculate the correlation coefficient between my independent and dependent variables as shown in Table 1. The correlation coefficient is around 0.3, which complements my analysis that there is a weak positive relationship between the proportion of seats taken by women in national parliaments and the labor force participation rate of females.


```{r}
## Finding correlation

corr <- cor(join_wdi_pol$prop_seats_women, join_wdi_pol$labor_female,
    use = "pairwise")
knitr::kable(corr, caption = "Correlation between proportion of seats taken by women in parliaments and female labor force participation", col.names = c("Correlation Coefficient"), digits = 2)
```



Furthermore, I plot the relationship between my two variables by regime type. I divide the data into three regime types upon their polity score. A polity score under -5 constitutes "authoritarian," a polity score of -5 to 5 constitutes an "anocracy," and a polity score above 5 constitutes a "democracy." Figure 3 shows the correlation between the proportion of seats held by women in parliament and the labor force participation of women by regime type. The plot for anocracies shows points scattered across the 0-30% range for women's participation in parliament and a wide range for female labor participation, with the exception of a few outliers that lie at a higher range for the former. The plot indicates a weak positive correlation between my two variables for anocracies. The plot for authoritarian regimes also shows that the points are widely dispersed, though none of the points lie above the 50% range for the proportion of seats held by women in parliament. However, there seems to be a stronger positive relationship between my two variables as compared to other regime types. Lastly, the plot for democracies shows the most randomness, as the points are scattered throughout all ranges of the proportion of seats taken by women in parliament and female labor force participation.


```{r}
## Plotting relationship by regime type
plot3 <- join_wdi_pol |>
  drop_na() |>
  mutate(regime_type = case_when(
    polity_score < -5 ~ "Authoritarian",
    polity_score >= -5 & polity_score < 6 ~ "Anocracy",
    polity_score > 5 ~ "Democracy"
  )) |>
  ggplot(aes(x = prop_seats_women, y = labor_female)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~regime_type) +
  labs(x = "Proportion of Women in National Parliament",
       y = "Labor Force Participation of Females",
       title = "Figure 3. Women in Politics and the Economy by Regime Type")
plot3
```


I also calculate the correlation coefficient for each regime type, as displayed in Table 2. The correlation between the proportion of seats held by women in national parliaments and female labor force participation rate is highest for authoritarian regimes at 0.47 and lowest for democracies at 0.18. The correlation coefficient is 0.21 for anocracies. Across all regime types, there seems to be a general positive relationship between my two variables. However, the findings counter my hypothesis that the relationship between women's political and economic participation will be weakest under authoritarian regimes. Contrary to my original hypothesis, the correlation between my two variables seems weakest in democracies. Additionally, this correlation does not necessarily indicate a causal relationship between the two variables, although an increase in one seems to correlate with an increase in the other to an extent.


```{r}
corr_regime <- join_wdi_pol |>
  drop_na() |>
  mutate(regime_type = case_when(
    polity_score < -5 ~ "Authoritarian",
    polity_score >= -5 & polity_score < 6 ~ "Anocracy",
    polity_score > 5 ~ "Democracy"
  )) |>
  group_by(regime_type) |>
  summarize(cor(prop_seats_women, labor_female))

knitr::kable(corr_regime, caption = "Correlation cefficient by regime type", col.names = c("Regime Type", "Correlation Coefficient"), digits = 2)
```


REGRESSION

I also evaluate the relationship between my two variables using multiple regressions. First, I run a regression of the rate of labor force participation of females (`labor_female`) as the dependent variable and the proportion of seats taken by women in national parliaments (`prop_seats_women`). The intercept coefficient is 47.4, indicating that when X = 0, meaning the proportion of seats held by women in parliaments is 0, the rate of female labor participation is 47.4, on average. The estimated coefficient on `prop_seats_women` is 0.47, indicating that for each percentage point increase in the proportion of seats held by women in parliaments, the female labor force participation rate increases by 0.47 percentage points, on average. When setting alpha to 0.05, this increase is statistically significant as the p-value for `prop_seats_women` coefficient is less than 0.001. However, the R-squared for this model is 0.092, which suggests that the model is not an appropriate fit for the data.


```{r}
## Running regressions

fit1 <- lm(labor_female ~ prop_seats_women, data = join_wdi_pol)
summary(fit1)

var_label <- c(
  "(Intercept)" = "Intercept",
  "prop_seats_women" = "Proportion of female seats in parliament"
)

modelsummary::modelsummary(fit1,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_map = var_label)
```


Second, I run a regression with the same dependent and independent variables, but with an additional covariate of polity score (`polity_score`). When holding the polity score constant, the intercept coefficient decreases from 47.4 to 42, indicating that when the proportion of seats held by women in parliaments is 0, and the polity score is 0, the rate of female labor participation is 42, on average. The estimated coefficient on `prop_seats_women` also decreases from 0.47 to 0.45, indicating that for each percentage point increase in the proportion of seats held by women in parliaments, the female labor force participation rate increases by 0.45 percentage points, on average, holding constant the polity score. When setting alpha to 0.05, the second model suggests that this increase is statistically significant as the p-value for `prop_seats_women` coefficient is less than 0.001. This multiple regression seems to be a better fit to the data, as the R-squared increases from 0.092 to 0.125, with an adjusted R-squared of 0.123. 

```{r}
fit2 <- lm(labor_female ~ prop_seats_women + polity_score, data = join_wdi_pol)
summary(fit2)

var_label <- c(
  "(Intercept)" = "Intercept",
  "prop_seats_women" = "Proportion of female seats in parliament",
  "polity_score" = "Polity score"
)

modelsummary::modelsummary(fit2,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_map = var_label)
```


I run a third regression with the same dependent and independent variables but with additional covariates of polity score (`polity_score`) and GDP per capita (`GDPpc`). When holding polity score and GDP per capita constant, the intercept coefficient decreases to 41.47, indicating that when the proportion of seats held by women in parliaments is 0, polity score is 0, and GDP per capita is 0, the rate of female labor participation is 41.47, on average. The estimated coefficient on `prop_seats_women` also decreases to around 0.35, indicating that for each percentage point increase in the proportion of seats held by women in parliaments, the female labor force participation rate increases by 0.35 percentage points, on average, holding constant the polity score and GDP per capita. When setting alpha to 0.05, the third model suggests that this increase is statistically significant as the p-value for `prop_seats_women` coefficient is less than 0.001. This multiple regression seems to be the best fit to the data, as the R-squared increases to 0.175, with an adjusted R-squared of 0.174. 

```{r}
fit3 <- lm(labor_female ~ prop_seats_women + polity_score + GDPpc, data = join_wdi_pol)
summary(fit3)

var_label <- c(
  "(Intercept)" = "Intercept",
  "prop_seats_women" = "Proportion of female seats in parliament",
  "polity_score" = "Polity score",
  "GDPpc" = "GDP per capita"
)

modelsummary::modelsummary(fit3,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           coef_map = var_label)
```

While the regression models suggest that an increase in the proportion of seats held by women in national parliaments may be associated with an increase in female labor force participation rates, it is difficult to establish that the coefficient represents a causal effect because of potential confounders that may not be controlled for in the proposed regressions. Moreover, the R-squared values for the regressions are fairly low, suggesting that the models are not fit for the data. However, it seems that the regression model befits the data as polity scores and GDP per capita are held fixed, suggesting that the two are potential confounders that require statistical control.


## Conclusion

Women's political participation has been emphasized as one of the key avenues to achieve gender equality within the international community, gender quotas being a means to achieve a descriptive level of women's inclusion in states' decision-making processes. Accordingly, democratic and non-democratic states alike have implemented gender quotas to increase women's political participation from the 1990s onward. However, there is little evidence to show that this increase in women's political participation has substantial effects on women's economic status. Hence, I test the relationship between women's political and economic participation using correlation and regression models with the proportion of seats held by women in national parliaments as my independent variable and the rate of female labor force participation as my dependent variable. Contrary to my initial hypothesis, there seems to be little effect of women's participation in parliament on the economic participation of women across all regime types. Furthermore, it seems that there is a higher correlation between the two variables for authoritarian states rather than democratic states. However, when holding polity score and GDP per capita fixed, a unit increase in the proportion of seats taken by women in parliaments is associated with a 0.35 percentage point increase in women's labor force participation. Such results question the substantial effects of women's inclusion in political institutions, even more so for democracies.

Nevertheless, my analysis is limited in two key ways. First, there is a lack of data both in terms of the dependent and independent variables, as well as polity scores. The problem of missing data is also more prevalent in authoritarian regimes, which presents a bias in our sample. States that have relatively higher female political and economic participation may have more incentives to report their statistics. Moreover, there may be potential confounders such as the educational attainment level of women, age demographics, and religion that may affect both the proportion of seats held by women in parliaments and the rate of female labor force participation. To advance the study further, more data on the missing variables such as polity score, women's labor force participation, and education level must be collected, especially for authoritarian regimes. Additionally, established gender quotas for each country may be added to the data set to examine a direct relationship between the quotas and women's economic participation.