---
title: "Women's Political and Economic Participation in Democracies and Autocracies"
author: "Han Na Jun"
description: "GOV50 Final Project"
output:
  distill::distill_article:
    self_contained: false
---


## Introduction

Do women's political participation in parliament positively affect women's economic participation? How do the correlation between these two factors vary across regime types, specifically democratic and authoritarian states? This study draws upon existing literature that investigates authoritarian regimes' incentives for implementing gender quotas, sometimes at higher levels than democratic states. While scholars have argued that authoritarian states do so to gain legitimacy at the domestic and international level, there is relatively little research as to whether such policies addressing women's political representation affect women's empowerment, specifically economic status. Accordingly, I seek to investigate the efficacy of women's heightened political participation under such regimes and discover if the policies go beyond mere signaling. I hypothesize that women's political participation in parliament will not have a significant effect on the economic status of women in authoritarian regimes. I also hypothesize that this tendency will be higher for more repressive regimes. This research contributes to the understanding of the substantial effects of women's inclusion in political institutions such as parliament in autocratic states, which is especially pertinent to the appeal for the implementation of gender quotas at the international level.


## Data Selection

My data draws upon the World Development Indicators Database produced by the World Bank and polity score data produced by the Center for Systemic Peace. The unit of analysis is by country and year, and the scope of my analysis is across 217 countries and from years 1990 to 2016, considering the rapid increase in gender quotas since the 1990s and the availability of data. My explanatory variable of interest is the 'Proportion of seats held by women in national parliaments (%),' which is a country-level measurement of women's political participation. My outcome variable of interest is the 'Labor force participation rate, female (% of female population ages 15-64),' which is a statistical model estimate by the International Labor Organization. Furthermore, I will compare these results by their relative polity score, which is a country-level coding of authority characteristics of states.

My research is a cross-sectional design that compares the outcome of the treatment, or political participation of women. If I observe little change in the labor force participation of women despite increase in the number of women in parliament under authoritarian regimes, this would provide support for my hypothesis. If I observe an increase in labor force participation of women along with the increase of women in parliament, the evidence would counter my hypothesis.

Figure 1 is a histogram that summarizes the labor force participation of females across all 217 countries from years 1990 to 2016. The histogram is skewed to the left, indicating that a significant portion of the data lies within 50% to 75% labor force participation rate of females.

```{r}
## Preparing the data

library(tidyverse)
wb_gender <- read_csv("data/wb_gender.csv")
polity <- read_csv("data/polity.csv")

wb_gender <- wb_gender |>
  select(`Country Name`, `Country Code`, `Series Name`, `1990 [YR1990]`:`2016 [YR2016]`)

colnames(wb_gender) <-c("country","ccode","series",1990:2016)

wdi <- wb_gender |>
  gather(key = year, value=value, `1990`:`2016`, factor_key = F) |>
  filter(series == "Proportion of seats held by women in national parliaments (%)" | series == "Labor force participation rate, female (% of female population ages 15-64) (modeled ILO estimate)" | series =="GDP per capita (constant 2015 US$)") |>
  pivot_wider(
    names_from = "series",
    values_from = "value"
  )

colnames(wdi) <- c("country", "ccode", "year", "prop_seats_women", "GDPpc", "labor_female")


wdi$year <- as.numeric(wdi$year)

head(wdi)
```

```{r}
##Joining WDI and Polity data

polity <- polity |>
  pivot_longer(cols = `AFG`:`ZIM`,
    names_to = "ccode",
    values_to = "polity_score"
  ) |>
  mutate(year = as.numeric(year))
polity

join_wdi_pol <- wdi |>
  left_join(polity, by = c("ccode", "year"))
join_wdi_pol

```


```{r}
## Plotting explanatory variable

plot1 <- wdi |>
  ggplot(aes(x = labor_female)) +
  geom_histogram() +
  labs(title = "Figure 1. Labor Force Participation of Females, 1990 - 2016",
       x = "Female labor force participation rate")
plot1
```


## Results

- For the results section, you should interpret (in plain English) the main coefficient of interest in your regression. You should also comment on the statistical significance of the estimated coefficient and whether or not you believe the coefficient to represent a causal effect.
2-3 paragraphs (8pts)

Figure 3 shows the correlation between the proportion of seats held by women in parliament and the labor force participation of women by regime type determined by the polity score. The graph indicates a weak positive correlation between the two variables across all types of regimes. This implies that women's political participation will positively correlate with women's economic participation in authoritarian regimes as well as in democracies. Thus, though a causal relationship between the two variables cannot be established, women's political participation may be a good indicator for women's economic participation and vice versa.


```{r}
## Plotting dependent and independent variables

plot2 <- join_wdi_pol |>
  ggplot(aes(x = prop_seats_women,
             y = labor_female)) +
  geom_point() +
  geom_smooth() +
  labs(title = "Figure 2. Political and Economic Participation of Females, 1990 - 2016",
       x = "Proportion of seats held by women in national parliaments (%)",
       y = "Female labor force participation rate")
plot2
```

```{r}
## Finding correlation

corr <- cor(join_wdi_pol$prop_seats_women, join_wdi_pol$labor_female,
    use = "pairwise")
corr
```

```{r}
## Plotting relationship by regime type
plot3 <- join_wdi_pol |>
  mutate(regime_type = case_when(
    polity_score < -5 ~ "Authoritarian",
    polity_score >= -5 & polity_score < 6 ~ "Anocracy",
    polity_score > 5 ~ "Democracy"
  )) |>
  ggplot(aes(x = prop_seats_women, y = labor_female)) +
  geom_point() +
  facet_wrap(~regime_type) +
  labs(x = "Proportion of Women in National Parliament",
       y = "Labor Force Participation of Females",
       title = "Figure 3. Women in Politics and the Economy by Regime Type")
plot3
```


```{r}
## Running regressions

fit1 <- lm(labor_female ~ prop_seats_women, data = join_wdi_pol)
summary(fit1)

fit2 <- lm(labor_female ~ prop_seats_women + polity_score, data = join_wdi_pol)
summary(fit2)

fit3 <- lm(labor_female ~ prop_seats_women + polity_score + GDPpc, data = join_wdi_pol)
summary(fit3)

fit4 <- lm(labor_female ~ GDPpc, data = join_wdi_pol)
summary(fit4)
```
regression explanation

## Conclusion

- a brief (one paragraph) concluding section that summarizes your results, assesses the extent to which you find support for your hypothesis, describes limitations of your analysis and threats to inference, and states how your analysis could be improved (e.g., improved data that would be useful to collect).
- 1 paragraph (i) summarizing results and assessing the extent to which you find support for your hypothesis; (ii) describing limitations of the analysis and threats to inference (missing data, confounding, etc), and stating how you could improve your analysis if you had more time/money. (2pts)


